---
title: "Assignment 6 Integer programming"
author: "Ananth Kumar"
date: "19/11/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
    geometry: margin=1in
---
library(rmarkdown)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F,message = F)
```
.libPaths("C:\\Users\\Ananth\\OneDrive\\Desktop\\MSBA Kent\\Fall 2021\\Fundamentals of Machine Learning\\Assignment\\Ass 2 ")


__Q1__ 
 
 
Decision variable:

Xij = 1, the arc from node i to node j is chosen in the optimal (longest) path otherwise Xij = 0
Objective Function:

Maximize the total time required from node 1 to node 9:
Max. Z = sigma(cij)(Xij)

Where, cij = time taken by arc (activity) from ith node and jth node

Max Z. = 5X12 + 3X13 + 3X35 + 2X25 + 4X24 + 4X47 + 1X46 + 2X58 + 6X57 + 5X69 + 4X79 + 7X89
Constraint:

For longest route problem, following constraint are to be satisfied,
For origin Node 1, outgoing arc is equal to 1,
X12 + X13 = 1

For intermediate nodes,
Arc in = arc out

For node 2: X12 = X25 + X24   or X12 – X25 - X24 = 0  
For node 3: X13 = X35   or X13 – X35 = 0
For node 4: X24 = X46 + X47   or X24 – X46 – X47 = 0
For node 5: X25 + X35 = X57 + X58 or X25 + X35 - X57 - X58 = 0
For node 6: X46 = X69   or X46 – X69 = 0
For node 7: X57 + X47 = X79   or X57 + X47 – X79 = 0
For node 8: X58 = X89   or X58 – X89 = 0

For destination node:
Arc in = 1
For node 9, X69 + X79 + X89 = 1


```{r}
library(lpSolve)
library(lpSolveAPI)
Q1 <- read.lp("Ques1.lp")

```

```{r}

solve(Q1)
get.objective(Q1) # longest path value is 17
get.variables(Q1) # each variable x12,x13,x35,x25,x24,x47,x46,x58,x57,x69,x79,x89 values, 1 indicates that the route is taken. 
```
LP equation for the problem 1 

// Objective function

max: 5 * x12 + 3 * x13 + 3 * x35 + 2 * x25 + 4 * x24 + 4 * x47 + x46 + 2 * x58 + 6 * x57 + 5 * x69 + 4 * x79 + 7 * x89;

// Constraints
 x12 + x13 = 1;
 x12 - x24 - x25 = 0;
 x13 - x35 = 0;
 x24 - x46 - x47 = 0;
 x25 + x35 - x57 - x58 = 0;
 x46 - x69 = 0;
 x57 + x47 - x79 = 0;
 x58 - x89 = 0;
 x69 + x79 + x89 = 1;

bin x12,x13,x35,x25,x24,x47,x46,x58,x57,x69,x79,x89;



__Q2__


__Selecting an Investment Portfolio__


The objective function should show maximum return to the amount invested based on the given constraints. 

/*Objective function*/
max: 4 s1 + 6.50 s2 + 5.9 s3 + 5.4 h1 + 5.15 h2 + 10 h3 + 8.4 c1 + 6.25 c2;



/*Constraints */ 

Limiting the total amount invested <= 2.5 million, 
40 s1 + 50 s2 + 80 s3 +60 h1 + 45 h2 + 60 h3 + 30 c1 + 25 c2 <= 2500000; 

Limiting the total amount invested per sector to <= 1 million , the stock price * number of shares of each company in a sector <= 1 million
40 s1 + 50 s2 + 80 s3 <= 1000000;
60 h1 + 45 h2 + 60 h3<= 1000000;
30 c1 + 25 c2 <= 1000000; 

Each company should atleat have 100,000 investment. the stock price * number of shares of a company <= 100,000
40 s1 >= 100000;
50 s2 >= 100000;
80 s3 >= 100000;
60 h1 >= 100000;
45 h2 >= 100000;
60 h3 >= 100000;
30 c1 >= 100000;
25 c2 >= 100000;

To get stock of each company which should be multiplier of 1000, we introduce a n1 which is an interger, which will make sure that the stock are multiples of 1000
s1 = 1000 n1;
s2 = 1000 n2;
s3 = 1000 n3;
h1 = 1000 n4;
h2 = 1000 n5;
h3 = 1000 n6;
c1 = 1000 n7;
c2 = 1000 n8;

Minimum stock should be 1, making sure that the amount is positive. We do not want any negative values.
n1 >= 1;
n2 >= 1;
n3 >= 1;
n4 >= 1;
n5 >= 1;
n6 >= 1;
n7 >= 1;
n8 >= 1;

n1 to n8 should be a positive integer to get share multiplier of 1000
int n1, n2, n3, n4, n5, n6, n7, n8;



```{r}

Question2P2 <-read.lp("Nointeger.lp")

solve(Question2P2)
get.objective(Question2P2)
get.variables(Question2P2)



```

_1) Determine the maximum return on the portfolio. What is the optimal number of shares to buy for each of the stocks? What is the corresponding dollar amount invested in each stock?__

__No integer restriction__

Total stocks owned in s1 = 4000 , Amount = 4000 * 40 = 160,000 $
Total stocks owned in s2 = 4000 , Amount = 4000 * 50 = 200,000 $   
Total stocks owned in s3 = 2000 , Amount = 2000 * 80 = 160,000 $
Total stocks owned in h1 = 2000 , Amount = 2000 * 60 = 120,000 $
Total stocks owned in h2 = 3000 , Amount = 3000 * 45 = 135,000 $
Total stocks owned in h3 = 12000, Amount = 12000 * 60= 720,000 $
Total stocks owned in c1 = 30000, Amount = 30000 * 30= 900,000 $
Total stocks owned in c2 = 4000 , Amount = 4000 * 25 = 100,000 $


Maximum return on portfolio:  477050

Return per share for the amount invested in s1 = 4 
Return per share for the amount invested in s2 = 6.5
Return per share for the amount invested in s3 = 5.9 
Return per share for the amount invested in h1 = 5.4 
Return per share for the amount invested in h2 = 5.15
Return per share for the amount invested in h3 = 10	 
Return per share for the amount invested in c1 = 8.4 
Return per share for the amount invested in c2 = 6.25


Maximum return = 4+6.5+5.9+5.4+5.15+10+8.4+6.25 = 51.6 (cumulative return including all stocks growth + dividend)


------------------------------------------------
------------------------------------------------
------------------------------------------------



Limiting the total amount invested <= 2.5 million, 
40 s1 + 50 s2 + 80 s3 +60 h1 + 45 h2 + 60 h3 + 30 c1 + 25 c2 <= 2500000; 

Limiting the total amount invested per sector to <= 1 million , the stock price * number of shares of each company in a sector <= 1 million
40 s1 + 50 s2 + 80 s3 <= 1000000;
60 h1 + 45 h2 + 60 h3<= 1000000;
30 c1 + 25 c2 <= 1000000; 

Each company should atleat have 100,000 investment. the stock price * number of shares of a company <= 100,000
40 s1 >= 100000;
50 s2 >= 100000;
80 s3 >= 100000;
60 h1 >= 100000;
45 h2 >= 100000;
60 h3 >= 100000;
30 c1 >= 100000;
25 c2 >= 100000;

To get stock of each company which should be multiplier of 1000, we introduce a n1 which is an interger, which will make sure that the stock are multiples of 1000
s1 = 1000 n1;
s2 = 1000 n2;
s3 = 1000 n3;
h1 = 1000 n4;
h2 = 1000 n5;
h3 = 1000 n6;
c1 = 1000 n7;
c2 = 1000 n8;

Minimum stock should be 1, making sure that the amount is positive. We do not want any negative values.
n1 >= 1;
n2 >= 1;
n3 >= 1;
n4 >= 1;
n5 >= 1;
n6 >= 1;
n7 >= 1;
n8 >= 1;

int s1, s2, s3, h1, h2, h3, c1, c2;

adding integer restriction for stocks:
int n1, n2, n3, n4, n5, n6, n7, n8; 



__Integer Restriction__



```{r}
integerrestriction <-read.lp("integerRestriction.lp")
solve(integerrestriction)
get.objective(integerrestriction)
get.variables(integerrestriction)
```





---------------------------------------------------------------------------------
 

__Compare the solution in which there is no integer restriction on the number of shares invested.__

Maximum Return : 473050
Total stocks owned in s1 = 3000 , Amount = 3000 * 40 = 120,000 $
Total stocks owned in s2 = 4000 , Amount = 4000 * 50 = 200,000 $   
Total stocks owned in s3 = 2000 , Amount = 2000 * 80 = 160,000 $
Total stocks owned in h1 = 2000 , Amount = 2000 * 60 = 120,000 $
Total stocks owned in h2 = 3000 , Amount = 3000 * 45 = 135,000 $
Total stocks owned in h3 = 12000, Amount = 12000 * 60= 720,000 $
Total stocks owned in c1 = 30000, Amount = 30000 * 30= 900,000 $
Total stocks owned in c2 = 4000 , Amount = 4000 * 25 = 100,000 $




___By how much (in percentage terms) do the integer restrictions alter the value of the optimal objective function?___

	             No Integer Restriction 	Integer Restriction 	 The Change	   Alter Percent 
	             
Optimum Value   	  477050	                  473050	            -4000	        -0.838%






__By how much (in percentage terms) do they alter the optimal investment quantities?__


	No Integer Restriction 	     Integer Restriction 	   The Change	    Alter Percent 
S1	1,60,000	                        1,20,000	         -40,000	         -25%
S2	2,00,000	                        2,00,000		
S3	1,60,000	                        1,60,000		
H1	1,20,000	                        1,20,000		
H2	1,35,000	                        1,35,000		
H3	7,20,000                         	7,20,000		
C1	9,00,000	                        9,00,000		
C2	1,00,000	                        1,00,000		
    ---------                         --------            -------         ------   
  	24,95,000	                        24,55,000	         -40,000	        -1.60%
 
 
 Alter Percent Aggregate : -1.60%


